Vod Report

### 2018-12-11 vodnotebook.Rmd
updated plotting and dst regressions

```{r}
cities_w_data_problems <- c("Green Bay", "Camden", "Bakersfield", "Little Rock")
```

## Load data & clean up

```{r}
data <- 
  read_rds(here::here("cache", "aggregated_cities_wsunset.Rds")) %>% 
  filter(!(city %in% cities_w_data_problems)) %>% 
  filter(minute < sunset_minute | minute > sunset_minute + 30)

subgeos <- 
  data %>% 
  group_by(city) %>% 
  summarise_at(
    vars(
      neighborhood, 
      region, 
      precinct, 
      reporting_area, 
      district, 
      beat, 
      sector, 
      police_grid_number
    ), 
    n_distinct,
    na.rm = TRUE
  ) %>% 
  gather(-city, key = "subgeo", value = "n_distinct") %>% 
  filter(
    n_distinct >= 5, n_distinct <= 30, 
    (city != "Dallas" | subgeo != "district")
  ) %>% 
  select(-n_distinct)

get_subgeo_cities <- function(name) {
  subgeos %>% 
    filter(subgeo == name) %>% 
    pull(city)
}

region_cities <- get_subgeo_cities("region")
precinct_cities <- get_subgeo_cities("precinct")
district_cities <- get_subgeo_cities("district")
beat_cities <- get_subgeo_cities("beat")
sector_cities <- get_subgeo_cities("sector")

data <-
  data %>% 
  filter(city %in% pull(subgeos, city)) %>% 
  mutate(
    subgeography = case_when(
      city %in% region_cities ~ region,
      city %in% precinct_cities ~ precinct,
      city %in% district_cities ~ district,
      city %in% beat_cities ~ beat,
      city %in% sector_cities ~ sector
    )
  ) %>% 
  filter(!is.na(subgeography)) %>% 
  unite("geo_adj_var", city, subgeography, remove = FALSE)
```

@@

```{r}
data %>% 
  ggplot(aes(time, fill = is_dark)) +
   geom_density(alpha = 1/2)
```

## Visualize

```{r}
plot_prop_minority_by_time(
  data, 
  title = "All stops", 
  min_time = hms::hms(hours = 18, min = 30), 
  max_time = hms::hms(hours = 19, min = 30)
)
```

```{r}
plot_prop_minority_by_time(
  data, 
  title = "All stops", 
  min_time = hms::hms(hours = 18, min = 30), 
  max_time = hms::hms(hours = 19, min = 30),
  smooth_method = "lm"
)
```

```{r}
data %>% filter(is_dst_period == TRUE) %>% count(city, sort = TRUE)
```

Use city with most stops around DST: Philadelphia

Get days of daylight savings:
```{r}
dst_dates <- 
  data %>% 
  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
  arrange(date) %>% 
  mutate(
    year = year(date),
    prev_dst = lead(is_dst),
    season = if_else(month < 6, "spring", "fall")
  ) %>% 
  filter(prev_dst != is_dst) %>% 
  select(year, dst_date = date, season)
```

Find time period  
```{r}
data %>% 
  group_by(date) %>% 
  summarise(sunset = min(sunset)) %>% 
  mutate(prev_sunset = lag(sunset)) %>% 
  filter(date %in% pull(dst_dates, dst_date)) %>% 
  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
  ggplot(aes(date, sunset_time, color = day, group = day)) +
  geom_point() +
  geom_line() +
  facet_grid(. ~ season)
```

5:15-5:30 in the fall and 6:10-6:25 in the spring

```{r}
data %>% 
  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
  mutate(
    year = year(date), 
    season = if_else(month > 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  mutate(days_since_dst = date - dst_date) %>%
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    season == "spring" & time < hms::hms(hours = 18, min = 25) & time > hms::hms(hours = 18, min = 10) |
      season == "fall" & time < hms::hms(hours = 17, min = 30) & time > hms::hms(hours = 17, min = 15)
  ) %>% 
  group_by(days_since_dst, season, is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
  geom_vline(xintercept = days(0)) +
  geom_point(aes(size = n), alpha = 1/2) +
  geom_smooth(se = FALSE, method = "lm") +
  facet_grid(. ~ season) +
  labs(
    title = "Philadelphia (2015-2018)",
    subtitle = "Time periods: 5:15-5:30 in the Fall & 6:10-6:25 in the Spring"
  )
```

Same for Nashville:
Get days of daylight savings:
```{r}
dst_dates_nashville <- 
  data %>% 
  filter(city == "Nashville", is_dst_period == TRUE) %>% 
  arrange(date) %>% 
  mutate(
    year = year(date),
    prev_dst = lead(is_dst),
    season = if_else(month < 6, "spring", "fall")
  ) %>% 
  filter(prev_dst != is_dst) %>% 
  select(year, dst_date = date, season)
```

```{r}
data %>% 
  group_by(date) %>% 
  summarise(sunset = min(sunset)) %>% 
  mutate(prev_sunset = lag(sunset)) %>% 
  filter(date %in% pull(dst_dates_nashville, dst_date)) %>% 
  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
  ggplot(aes(date, sunset_time, color = day, group = day)) +
  geom_point() +
  geom_line() +
  facet_grid(. ~ season)
```
Filter to 2011 - 2014: 5:30-5:45 fall, 6:10-6:25 spring
```{r}
data %>% 
  filter(city == "Nashville", is_dst_period == TRUE, year(date) <= 2014, year(date) >= 2011) %>% 
  mutate(
    year = year(date), 
    season = if_else(month > 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  mutate(days_since_dst = date - dst_date) %>%
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    season == "spring" & time < hms::hms(hours = 18, min = 45) & time > hms::hms(hours = 18, min = 30) |
      season == "fall" & time < hms::hms(hours = 17, min = 45) & time > hms::hms(hours = 17, min = 30)
  ) %>% 
  group_by(days_since_dst, season, is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
  geom_vline(xintercept = days(0)) +
  geom_point(aes(size = n), alpha = 1/2) +
  geom_smooth(se = FALSE, method = "lm") +
  facet_grid(. ~ season) +
  labs(
    title = "Nashville (2011-2014)",
    subtitle = "Time periods: 5:30-5:45 in the Fall; 6:30-6:45 in the Spring"
  )
```

@@

### All data

```{r}
log(calculate_k(data))
```

```{r}
simple_model <-
  data %>% 
  group_by(is_dark, minute) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")

broom::tidy(simple_model)
```

```{r}
geo_model <-
  data %>% 
  group_by(is_dark, minute, geo_adj_var) %>% 
  summarise(
    n = n(), 
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

broom::tidy(geo_model)
```

### DST 
```{r}
dst_models <- read_rds(here::here("cache", "vod_dst_models.Rds"))
dst_results <- read_rds(here::here("cache", "vod_dst_results.Rds"))

dst_models$model_district_adjusted %>% broom::tidy()
```

```{r}
log(calculate_k(data %>% filter(is_dst_period == TRUE)))
```

```{r}
dst_simple_model <-
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  group_by(is_dark, minute) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")

broom::tidy(dst_simple_model)
```

```{r}
dst_geo_model <-
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  group_by(is_dark, minute, geo_adj_var) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

broom::tidy(dst_geo_model)
```

### Separately on each city

```{r}
reg <- function(city_name){
  model <-
    data %>% 
    filter(city == city_name) %>% 
    group_by(is_dark, minute, geo_adj_var) %>% 
    summarise(
      n = n(), 
      k = sum(is_minority_demographic)
    ) %>% 
    glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

  broom::tidy(model) %>% 
    filter(term == "is_darkTRUE") %>% 
    mutate(city = city_name) %>% 
    select(city, estimate, std.error)
}

by_city <- 
  subgeos %>% 
  pull(city) %>% 
  map_dfr(~reg(.))

by_city %>% 
  filter(city != "Plano") %>% 
  mutate(low = estimate - std.error, high = estimate + std.error) %>% 
  ggplot(aes(reorder(city, estimate), estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high))
```

>>>>

## Load data

```{r}
data <- read_rds(here::here("cache", "vod_full_data.Rds"))
```

### summary table

```{r}
data %>% 
  group_by(city) %>% 
  summarise(
    n = n(),
    n_dst = sum(is_dst_period),
    proportion_minority = sum(is_minority_demographic) / n
  ) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```

@@

```{r}
data %>% 
  filter(city == "Philadelphia") %>% 
  ggplot(aes(time, fill = is_dark)) +
   geom_density(alpha = 1/2)
```

## Visualize All Stops

```{r}
plot_prop_minority_by_time(
  data, 
  title = "All stops", 
  min_clock_time = hms::hms(hours = 18, min = 30), 
  max_clock_time = hms::hms(hours = 19, min = 30)
)
```

```{r}
plot_prop_minority_by_time(
  filter(data, city != "El Paso"),
  min_clock_time = hms::hms(hours = 18, min = 30), 
  max_clock_time = hms::hms(hours = 19, min = 30),
  smooth_method = "lm"
)
```

```{r}
plot_prop_minority_by_time(
    data,
    city_only = "Madison",
    title = "Madison, WI",
    min_clock_time = hms::hms(hours = 18, min = 30), 
    max_clock_time = hms::hms(hours = 19, min = 30),
    smooth_method = "lm"
  )
```

### Regression Results

#### Summary
```{r}
full_results
```

#### Time adjusted only

#### Time and subgeography adjusted

```{r}
full_models$model_geo_adjusted %>% broom::tidy()
```

## Daylight Savings Period

@@

```{r}
dst_fall <- read_rds(here::here("cache", "vod_dst_fall.Rds"))
dst_spring <- read_rds(here::here("cache", "vod_dst_spring.Rds"))
```

#### Summary
```{r}
tribble(
  ~adjustments, ~Fall_logK,  ~Fall_SE, ~Spring_logK,  ~Spring_SE,
  "None",       log(dst_fall$k), NA, log(dst_spring$k),  NA,
  "Clock time", -coef(dst_fall$model_time_const)[2], coef(summary(dst_fall$model_time_const))[2, 2], -coef(dst_spring$model_time_const)[2], coef(summary(dst_spring$model_time_const))[2, 2],
  "Clock time and Subgeography", -coef(dst_fall$model_geo_adjusted)[2], coef(summary(dst_fall$model_geo_adjusted))[2,2], -coef(dst_spring$model_geo_adjusted)[2], coef(summary(dst_spring$model_geo_adjusted))[2,2]
)
```

## Separately on each city

```{r}
by_city <- read_rds(here::here("cache", "vod_geo_adj_by_city.Rds"))

by_city %>%  
  mutate(
    log_k = -estimate, 
    low = log_k - 1.96*std.error, 
    high = log_k + 1.96*std.error
  ) %>% 
  ggplot(aes(reorder(city, log_k), log_k)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high)) +
  theme(
    axis.text.x = element_text(hjust = 1, angle = 30)
  ) +
  labs(
    x = NULL,
    y = "logK with 95% conf int",
    title = "Clock-time, Month and Subgeography Adjusted Models by City"
  )
```

```{r}
by_city_dst <- read_rds(here::here("cache", "vod_dst_by_city.Rds"))

by_city_dst %>%  
  mutate(low = estimate - 1.96*std.error, high = estimate + 1.96*std.error) %>% 
  ggplot(aes(reorder(city, estimate), estimate)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high)) +
  theme(
    axis.text.x = element_text(hjust = 1, angle = 30)
  ) +
  labs(
    x = NULL,
    y = "Coefficient on isDarkTRUE with 95% conf int",
    title = "Clock-time and Subgeography Adjusted DST Models by City"
  ) +
  facet_grid(season ~ .)
```

## discontinuity

```{r}
date_lims <- 
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  summarise(min = min(date), max = max(date))

city_years <- 
  data %>% 
  distinct(city, year = year(date))

city_centroids <-
  data %>% 
  group_by(city) %>% 
  summarise(
    lat = median(lat),
    lng = median(lng),
    tz = min(tz)
  )

dst_sunsets <-
  tibble(
     date = seq(date_lims$min, date_lims$max, by = "days")
  ) %>% 
  mutate(
    is_dst = dst(as.POSIXct(date, tz = Sys.timezone())),
    is_dst_tomorrow = lead(is_dst),
    day_of_change = is_dst == FALSE & is_dst_tomorrow == TRUE | 
      is_dst == TRUE & is_dst_tomorrow == FALSE,
    day_before_change = lead(day_of_change) == TRUE
  ) %>% 
  filter(day_of_change == TRUE | day_before_change == TRUE) %>% 
  mutate(
    year = year(date),
    season = if_else(month(date) < 6, "fall", "spring")
  ) %>% 
  select(year, season, date, day_before_change, day_of_change) %>% 
  left_join(city_years, by = "year") %>% 
  left_join(city_centroids, by = "city") %>% 
  mutate(subgeography = city) %>% 
  left_join(infer_sunset_times(.), by = c("date", "tz", "subgeography"))

dst_times <-
  dst_sunsets %>% 
  mutate(
    day = if_else(day_before_change == TRUE, "day_before", "day_of"),
    sunset_minute = hour(hms(sunset)) * 60 + minute(hms(sunset))
  ) %>% 
  select(-day_before_change, -day_of_change, -date, -sunset, -dusk) %>% 
  spread(key = day, value = sunset_minute) %>% 
  mutate(
    earlier = if_else(day_before < day_of, day_before, day_of),
    later = if_else(day_before > day_of, day_before, day_of),
    min_minute = earlier + 20,
    max_minute = earlier + 40
  ) %>% 
  filter(max_minute < later) %>% 
  select(year, season, city, min_minute)

dst_dates <-
  dst_sunsets %>% 
  filter(day_of_change) %>% 
  distinct(year, season, date) %>% 
  select(year, season, dst_date = date)

full_data <-
  data %>% 
  mutate(
    year = year(date),
    season = if_else(month < 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  left_join(dst_times, by = c("year", "season", "city")) %>% 
  mutate(days_since_dst = date - dst_date)

plot_data <-
  full_data %>% 
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    minute >= min_minute,
    minute <= min_minute + 20,
    city != "El Paso"
  ) %>% 
  mutate(
    is_dark = factor(is_dark, levels = c(TRUE, FALSE), labels = c("Dark", "Light")),
    before_dst = days_since_dst < 0
  ) %>% 
  group_by(days_since_dst, before_dst, season, is_dark, geo_control) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>%
  summarise(weighted_mean_prop = weighted.mean(prop_minority, n))

plot_data %>% 
  ggplot(aes(days_since_dst, weighted_mean_prop)) +
  geom_vline(xintercept = days(0), color = "grey60") +
  geom_point(aes(color = is_dark)) +
  geom_smooth(aes(group = is_dark), method = "lm", color = "grey40", se = FALSE) +
  scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
  labs(
    title = "DST Discontinuity (20 minute period) over 9 cities",
    x = "Days since DST change",
    y = "Proportion minority drivers stopped - Mean over cities",
    color = NULL
  ) +
  facet_grid(season ~ .)
```

## Statewide

Choice of states: race & time & county available. chose to override recorded lat/lng (partial in CT, TX, WI; good coverage MT, OH, VT) and drop VT which has only lat/lng no county due to lat/lng errors & for consistency.

```{r}
statewide_results <- read_rds(here::here("cache", "vod_statewide_full_results.Rds"))
statewide_models <- 
  list(
    time_only = read_rds(here::here("cache", "vod_statewide_time_const.rds")),
    time_geo = read_rds(here::here("cache", "vod_statewide_geo_adj.rds"))
  )

tribble(
  ~adjustments, ~logK,                                    ~Standard_Error,
  "None",       statewide_results$log_k,                       NA,
  "Clock time", -statewide_results$coefficients$time_adjusted, statewide_models$time_only$std.error[2],
  "Clock time and Subgeography", -statewide_results$coefficients$time_and_geo_adjusted, statewide_models$time_geo$std.error[2]
)
```

```{r}
statewide_dst_results <- read_rds(here::here("cache", "vod_statewide_dst_results.Rds"))
statewide_dst_models <- read_rds(here::here("cache", "vod_statewide_dst_models.Rds"))

tribble(
  ~adjustments, ~logK,                                    ~Standard_Error,
  "None",       statewide_dst_results$log_k,                       NA,
  "Clock time", -statewide_dst_results$coefficients$time_adjusted, coef(summary(statewide_dst_models$model_time_const))[2, 2],
  "Clock time and Subgeography", -statewide_dst_results$coefficients$time_and_geo_adjusted, coef(summary(statewide_dst_models$model_geo_adjusted))[2,2]
)
```

```{r}
data_states %>% 
  ggplot(aes(time, fill = is_dark)) +
   geom_density(alpha = 1/2)
```

```{r}
discontinuity_states <- read_rds(here::here("cache", "states_dst_discontinuity.rds"))

discontinuity_states %>% 
  group_by(days_since_dst, season, is_dark, state) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n()) %>% 
  summarise(med_prop_minority = median(prop_minority)) %>% 
  ggplot(aes(days_since_dst, med_prop_minority)) +
  geom_vline(xintercept = days(0), color = "grey60") +
  geom_point(aes(color = is_dark)) +
  geom_smooth(aes(group = is_dark), se = FALSE, method = "lm", color = "grey20") +
  scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
  facet_grid(. ~ season) +
  labs(
    title = "DST Discontinuity (20 minute period) over 10 states",
    x = "Days since DST change",
    y = "Proportion minority drivers stopped - Mean over states",
    color = NULL
  )
```

```{r}
city %>%

```
